\bf Calendar-based graphics for visualizing people's daily schedules

### 2019-04-23 reviews/response.Rmd

Thanks to the three reviewers for some very helpful comments which have led to many improvements of the paper. My response are in black, the reviewers' comments are in red.

# Reviewer 1 comments {-}

> Figure 2 and Figure 10 show much the same information. You can see the Australian Open period in both. There is more detail in Figure 10 for the State Library.  The gaps in the Birrarung Marr data are easier to see in Figure 2.

We add a new subsetion "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information revealed by calendar plots is but not revealed by other plots. The calendar display doesn't replace traditional displays, but serves to complement to further tease out interesting structures in the data.

@@

> Figure 11 is not described well.  What does the V-pattern mean?  How can you see the strong autocorrelation at weekends?  There are only dots.

We have switched to use the local scaling for the lag scatterplot (as Figure 10 now) to accurately represent the autocorrelation. The strong autocorrelation is clearly revealed at weekends.

@@

> The colours are not well chosen in Figure 15.  It is hard to see the averages for households 1 and 3.

Done.

> There are too many large plots in the paper.  Figure 6 is not needed and Figure 12 is not needed.  How do you know that the New Year's Eve outliers in Figure 12 are from the same sensor?  That is not in the plot.

Thanks for correcting it. We have pointed out the variability instead of the outliers on New Year's Eve.

@@

> Fig.2: Can you match the minor grid lines with the start of a month, rather than the middle of a month? This will make it much easier to identify the approximate dates for some of the other spikes. No need to label these minor grid lines.

Done.

> Fig.2: You mention "small multiples" only in a figure caption. This concept is a central
part of your calendar graphics and should be summarized in more details (including references) in the main text. Also mention Unwin & Valero-Mora's "Ensemble Graphics", JCGS, 27(1), as a major concept that applies to both sets of related figures in your article.

We have explained the idea of faceting and also referred to small multiples, trellis charts, and ensemble graphics at the end of the second paragraph in Introduction section.

@@

> Fig.8: "impossible to compare the size of peaks between days." - Correct; since this is based on ggplot, could you use colors from a sequential color scheme that maps the counts, e.g., from faint yellow (0) to dark red (for the overall max)? This might be another
useful argument for a future extension of your function.

ToDo.

> p15, l12: "idea of faceting": explain what "faceting" means in general (and cite 1 or 2 main references)

We added the explanation of "faceting" in the Introduction section.

### 2019-04-23 main.Rmd
new subsetion "reasons to use calendar-based graphics"

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. [ToDo: summarise small multiples in more details including citations.] The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series. The proposed algorithm has been implemented in the `frame_calendar` function in the **sugrrants** package using R.

@@

```{r boxplot, fig.height = 8, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day. It shows the hourly distribution in the city as a whole. There is one sensor attracting a larger number of people on New Year's Eve than the rest."}
```

@@

## Reasons to use calendar-based displays

>>>>

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \@ref(fig:ped-map)). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \@ref(fig:time-series-plot) give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \@ref(fig:facet-time)) provides a better glimpse of the daily and sub-daily pedestrian patterns. Faceting generates small multiples [@tufte1983visual] or trellis displays [@becker_visual_1996], where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets. A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics [@unwin_ensemble_2018].

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers. The proposed algorithm has been implemented in the `frame_calendar` function in the **sugrrants** package using R.

@@

```{r boxplot, fig.height = 8, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day. It shows the hourly distribution in the city as a whole. The increased varibility is noted on the last day of December as approaching towards New Year's Eve."}
```

@@

## Reasons to use calendar-based graphics

The calendar-based graphics provides a contextual canvas to plot glyphs into a familiar calendar format. The purpose of this calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is critical in data visualisation. It not only complements the traditional graphical toolbox for understanding general trends, but it profiles vivid and detailed data stories about the way we live. Comparing to Figure \@ref(fig:time-series-plot) and \@ref(fig:facet-time), Figure \@ref(fig:facet) is more informatively compelling: whether special events happened on which day and whether they are day or night events. Figure \@ref(fig:facet) informs that many events were held in Birrarung Marr during weekends, while September's events took place on Fridays, which is difficult to discern in the other two plots.

### 2019-04-23 main.tex
new subsetion "reasons to use calendar-based graphics"

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. {[}ToDo: summarise small multiples in more details including citations.{]} The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable overlaying plots consisting of multiple time series. The proposed algorithm has been implemented in the \texttt{frame\_calendar} function in the \textbf{sugrrants} package using R.

@@

\caption{Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day. It shows the hourly distribution in the city as a whole. There is one sensor attracting a larger number of people on New Year's Eve than the rest.}\label{fig:boxplot}
\end{figure}

@@

\hypertarget{reasons-to-use-calendar-based-displays}{%
\subsection{Reasons to use calendar-based displays}\label{reasons-to-use-calendar-based-displays}}

>>>>

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia \citep{ped}. There have been 43 sensors installed that count pedestrians every hour across the inner-city area until the end of 2016 (Figure \ref{fig:ped-map}). The data set can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot} give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns. Faceting generates small multiples \citep{tufte1983visual} or trellis displays \citep{becker_visual_1996}, where different subsets of the same data are displayed based on conditioning variables, allowing for comparison across subsets. A series of graphics can be further grouped together to create multiple yet different data views, that forms ensemble graphics \citep{unwin_ensemble_2018}.

@@

We propose a new algorithm to go beyond the calendar-based heatmap. The approach is developed with three conditions in mind: (1) to display time-of-day variation in addition to longer temporal components such as day-of-week and day-of-year; (2) to incorporate line graphs and other types of glyphs into the graphical toolkit for the calendar layout; (3) to enable patterns related to special events more easily pop-up to viewers. The proposed algorithm has been implemented in the \texttt{frame\_calendar} function in the \textbf{sugrrants} package using R.

@@

\caption{Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day. It shows the hourly distribution in the city as a whole. The increased varibility is noted on the last day of December as approaching towards New Year's Eve.}\label{fig:boxplot}
\end{figure}

@@

\hypertarget{reasons-to-use-calendar-based-graphics}{%
\subsection{Reasons to use calendar-based graphics}\label{reasons-to-use-calendar-based-graphics}}

The calendar-based graphics provides a contextual canvas to plot glyphs into a familiar calendar format. The purpose of this calendar display is to facilitate quick discoveries of unusual patterns in people's activities, which is critical in data visualisation. It not only complements the traditional graphical toolbox for understanding general trends, but it profiles vivid and detailed data stories about the way we live. Comparing to Figure \ref{fig:time-series-plot} and \ref{fig:facet-time}, Figure \ref{fig:facet} is more informatively compelling: whether special events happened on which day and whether they are day or night events. Figure \ref{fig:facet} informs that many events were held in Birrarung Marr during weekends, while September's events took place on Fridays, which is difficult to discern in the other two plots.
